### 2020-12-06 paper/localpsf.tex
more edits to intro

Our approximation is defined by how we compute the convolution kernels, $\varphi_i$, (Section \ref{sec:get_impulse_response}) how we choose the points, $x_i$, (Section \ref{sec:choosing_sample_points}) and what weighting functions, $w_i$, that we use (Section \ref{sec:weighting_functions}). 
Briefly, we compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources, $\delta_{x_i},$ at many points $x_i$ (a ``Dirac comb''). That is, we compute
\begin{equation*}
	\eta := \sum_{i=1}^m A(~\cdot~, x_i) = \left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle^*.
\end{equation*}
Here $\sum_{i=1}^m \delta_{x_i}$ is the Dirac comb distribution, and $\left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle^*$ denotes the Riesz representation of the application of $\mathcal{A}$ to the Dirac comb. In computations, this amounts to computing $\boldsymbol{\eta} = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \left(\sum_{i=1}^m \boldsymbol{\delta}_{x_i}\right)$, where $\boldsymbol{\eta}$, $\mathbf{A}$, and $\boldsymbol{\delta}_{x_i}$ are discretized versions of $\eta$, $A$, and $\delta_{x_i}$, respectively, and $\mathbf{M}$ is a mass matrix. The batch of points $x_i$ are chosen so that the support of $\left\langle \mathcal{A}, \delta_{x_i}\right\rangle^*$ and the support of $\left\langle \mathcal{A}, \delta_{x_j}\right\rangle^*$ do not overlap (or do not overlap much) if $i \neq j$. Since the impulse responses do not overlap, we may post-process $\eta$ to recover $\varphi_i$ for all of the points $x_i$ in the batch---by applying $\mathcal{A}$ to one distribution, we recover many $\varphi_i$. This is illustrated in Figure FIG. The process is repeated to get more batches of $\varphi_i$'s, until the approximation is sufficiently accurate or a maximum number of sample points is reached. In order to choose the points $x_i$, we need to estimate the support of $\left\langle \mathcal{A}, \delta_{x_i}\right\rangle^*$ \emph{before} we compute it. The supports of the impulse responses $\left\langle \mathcal{A}, \delta_{x}\right\rangle^*$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions (Section \ref{eq:mean_and_covariance_estimation}). This procedure is the only part of the paper that requires non-negativity of the integral kernel. We choose the functions $w_i$ to be the smoothest functions, in a least-squares sense, that interpolate the points $x_i$. Computing the functions $w_i$ require solving two Poisson PDEs per weighting function, and these PDEs are solved cheaply using multigrid.

@@

@@

@@

In this paper we present an efficient matrix-free method for approximating operators $\mathcal{A}$ with a product-convolution approximation, $\widetilde{A}$, of the form
\begin{equation*}
	\left(\mathcal{A}u\right) \approx \left(\widetilde{\mathcal{A}}u\right) := \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right)
\end{equation*}

PSFs that satisfy the following properties:
\begin{enumerate}
	\item $\varphi_x$ varies smoothly as a function of $x$.
	\item $\varphi_x(y)$ is zero (or small) when $|y|$ is large.
	\item $\varphi_x$ is a non-negative (or mostly non-negative) function.
\end{enumerate}
If $\varphi_x$ was constant in $x$, then $\mathcal{A}$ would be the convolution operator $\left(\mathcal{A}v\right)^* = \varphi_0 \ast v$. Operators satisfying property 1 therefore act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies smoothly as $x$ changes. Operators satisfying property 2 are local, in the sense that sources at point $x$ have no impact on points far away from $x$. Operators satisfying property 3 blur inputs: their action on a point source consists of scaling the source, then spreading the energy of the scaled point source across a region nonuniformly. We call operators satisfying these properties \emph{smoothly varying blur operators}.

@@

The functions $w_i$ are the defined as the smoothest functions, in a certain least-squares sense, that interpolate the points $x_i$. Computing the functions $w_i$ requires solving two Poisson PDEs for each point $x_i$, and these Poisson PDEs may be solved cheaply with multigrid.

The idea of our method is to sample $\varphi_{x}$ at a collection of points $x_i$ by applying $\mathcal{A}$ to a sum of point sources at the points $x_i$ (a ``Dirac comb''). The points $x_i$ are chosen so that the support of $\langle\mathcal{A},\delta_{x_i}\rangle^*$ and the support of $\langle\mathcal{A},\delta_{x_j}\rangle^*$ do not overlap (or do not overlap much) if $i \neq j$. Since the point sources do not overlap, by applying $\mathcal{A}$ to one vector, we recover $\varphi_{x_i}$ for many points $x_i$. The process is repeated to get $\varphi_{x_i}$ at more points, until an error estimate is satisfied. The $\varphi_{x_i}$ are interpolated smoothly to construct $\varphi_x$ for arbitrary $x$, which yields an approximation $\widetilde{A}(y,x)$ to the kernel $A(y,x)$, and an approximation $\widetilde{\mathcal{A}}$ to the operator $\mathcal{A}$. While $A(y,x)$ is inaccessible, we can evaluate $\widetilde{A}(y,x)$ at arbitrary $(y,x)$ points. This allows us to construct a hierarchical matrix representation of $\widetilde{\mathcal{A}}$, which allows us to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations, using fast hierarchical matrix arithmetic.

Properties 1 and 2 are essential to our method. Property 3 is only required in the subprocedure for estimating the support of $\langle\mathcal{A},\delta_x\rangle^*$, and may be dropped if other methods for estimating the support are available.

>>>>

\subsection{Overview of the method}

Our approximation method is defined by how we compute the convolution kernels, $\varphi_i$, (Section \ref{sec:get_impulse_response}) how we choose the points, $x_i$, (Section \ref{sec:choosing_sample_points}) and what weighting functions, $w_i$, that we use (Section \ref{sec:weighting_functions}). 
Briefly, we compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources, $\delta_{x_i},$ at many points $x_i$ (a ``Dirac comb''). That is, we compute
\begin{equation*}
	\eta := \sum_{i=1}^m A(~\cdot~, x_i) = \left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle^*.
\end{equation*}
Here $\sum_{i=1}^m \delta_{x_i}$ is the Dirac comb distribution, and $\left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle^*$ denotes the Riesz representation of the application of $\mathcal{A}$ to the Dirac comb. In computations, this amounts to computing $\boldsymbol{\eta} = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \left(\sum_{i=1}^m \boldsymbol{\delta}_{x_i}\right)$, where $\boldsymbol{\eta}$, $\mathbf{A}$, and $\boldsymbol{\delta}_{x_i}$ are discretized versions of $\eta$, $A$, and $\delta_{x_i}$, respectively, and $\mathbf{M}$ is a mass matrix. The batch of points $x_i$ are chosen so that the support of $\left\langle \mathcal{A}, \delta_{x_i}\right\rangle^*$ and the support of $\left\langle \mathcal{A}, \delta_{x_j}\right\rangle^*$ do not overlap (or do not overlap much) if $i \neq j$. Since the impulse responses do not overlap, we may post-process $\eta$ to recover $\varphi_i$ for all of the points $x_i$ in the batch---by applying $\mathcal{A}$ to one distribution, we recover many $\varphi_i$. This is illustrated in Figure FIG. The process is repeated to get more batches of $\varphi_i$'s, until the approximation is sufficiently accurate or a maximum number of sample points is reached. In order to choose the points $x_i$, we need to estimate the support of $\left\langle \mathcal{A}, \delta_{x_i}\right\rangle^*$ \emph{before} we compute it. The supports of the impulse responses $\left\langle \mathcal{A}, \delta_{x}\right\rangle^*$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions (Section \ref{eq:mean_and_covariance_estimation}). This procedure is the only part of the paper that requires non-negativity of the integral kernel.

We choose the weighting functions $w_i$ to be the smoothest functions, in a least-squares sense, that interpolate the points $x_i$. Computing the weighting functions $w_i$ requires solving two Poisson PDEs per weighting function, and these PDEs are solved cheaply using multigrid.

@@

\subsection{Connection to existing methods}

@@

\subsection{old stuff}

@@

%In this paper we present an efficient matrix-free method for approximating operators $\mathcal{A}$ with a product-convolution approximation, $\widetilde{A}$, of the form
%\begin{equation*}
%	\left(\mathcal{A}u\right) \approx \left(\widetilde{\mathcal{A}}u\right) := \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right)
%\end{equation*}
%
%
%PSFs that satisfy the following properties:
%\begin{enumerate}
%	\item $\varphi_x$ varies smoothly as a function of $x$.
%	\item $\varphi_x(y)$ is zero (or small) when $|y|$ is large.
%	\item $\varphi_x$ is a non-negative (or mostly non-negative) function.
%\end{enumerate}
%If $\varphi_x$ was constant in $x$, then $\mathcal{A}$ would be the convolution operator $\left(\mathcal{A}v\right)^* = \varphi_0 \ast v$. Operators satisfying property 1 therefore act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies smoothly as $x$ changes. Operators satisfying property 2 are local, in the sense that sources at point $x$ have no impact on points far away from $x$. Operators satisfying property 3 blur inputs: their action on a point source consists of scaling the source, then spreading the energy of the scaled point source across a region nonuniformly. We call operators satisfying these properties \emph{smoothly varying blur operators}.

@@

%The functions $w_i$ are the defined as the smoothest functions, in a certain least-squares sense, that interpolate the points $x_i$. Computing the functions $w_i$ requires solving two Poisson PDEs for each point $x_i$, and these Poisson PDEs may be solved cheaply with multigrid.

%The idea of our method is to sample $\varphi_{x}$ at a collection of points $x_i$ by applying $\mathcal{A}$ to a sum of point sources at the points $x_i$ (a ``Dirac comb''). The points $x_i$ are chosen so that the support of $\langle\mathcal{A},\delta_{x_i}\rangle^*$ and the support of $\langle\mathcal{A},\delta_{x_j}\rangle^*$ do not overlap (or do not overlap much) if $i \neq j$. Since the point sources do not overlap, by applying $\mathcal{A}$ to one vector, we recover $\varphi_{x_i}$ for many points $x_i$. The process is repeated to get $\varphi_{x_i}$ at more points, until an error estimate is satisfied. The $\varphi_{x_i}$ are interpolated smoothly to construct $\varphi_x$ for arbitrary $x$, which yields an approximation $\widetilde{A}(y,x)$ to the kernel $A(y,x)$, and an approximation $\widetilde{\mathcal{A}}$ to the operator $\mathcal{A}$. While $A(y,x)$ is inaccessible, we can evaluate $\widetilde{A}(y,x)$ at arbitrary $(y,x)$ points. This allows us to construct a hierarchical matrix representation of $\widetilde{\mathcal{A}}$, which allows us to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations, using fast hierarchical matrix arithmetic.

%Properties 1 and 2 are essential to our method. Property 3 is only required in the subprocedure for estimating the support of $\langle\mathcal{A},\delta_x\rangle^*$, and may be dropped if other methods for estimating the support are available.
